Quantifying fish range shifts across poorly defined management boundaries

### 2021-10-27 Scripts/manuscript_acrossboundaries.Rmd
FIrst discussion draft

Next, we estimated the CPUE percentage change today relative to the period used to determine the current quota for each stock. Today is defined as the average of 2009 to 2019 representing the more recent data available.

For summer flounder, the quota allocation period covers 1980 to 1986 determined in the original management plan [@Ammendment4; 9.1.2.3.1. Commercial quota page 12], for scup the period covers 1988 to 1992 according to Amendment 8 of the original management plan ([@Ammendment8]; **Section 9.1.2.3.7. Commercial quota, page 48 and Table 40**) and for black sea bass Amendment 13 sets the state-wide quota based on historic landings from 1980 to 2001 (@REF; **Amendment 13- Alternative 5f; section 2.1.5.6 (Pages 3, 14, and 304-Table 5**). The parentage changes between the historic and present CPUE ($\Delta{B}$) was estimate per stock for each of the two regulatory units approaches as;

@@

>>>>

Next, we estimated the CPUE percentage change today relative to a historic time period. We defined today as the average of 2009 to 2019 representing the more recent data available. The historic time period is determine by the years used to set the state-wide quota for each stock. For summer flounder, the quota allocation period covers 1980 to 1986 determined in the original management plan [@Ammendment4; 9.1.2.3.1. Commercial quota page 12], for scup the period covers 1988 to 1992 according to Amendment 8 of the original management plan ([@Ammendment8]; **Section 9.1.2.3.7. Commercial quota, page 48 and Table 40**) and for black sea bass Amendment 13 sets the state-wide quota based on historic landings from 1980 to 2001 (@REF; **Amendment 13- Alternative 5f; section 2.1.5.6 (Pages 3, 14, and 304-Table 5**). The parentage changes between the historic and present CPUE ($\Delta{B}$) was estimate per stock for each of the two regulatory units approaches as;

@@

Marine species have been shifting globally in response to climate change [@Bell; @Pinsky] and are expected to continue towards the end of the century [@Morey; @Pinsky]. Such shifts challenge ocean governance, particularly when dealing with shared stocks that cross international and national management boundaries [@PalaciosAbrantes2020; @Pisnky2018; @Oremus2020 @Pinsky2014; @Pinsky2014]. Managers are having to design new management strategies as is the case presented here of the Mid-Atlantic Fishery Management Council (MAFMC) who oversees black sea bass, summer flounder, and scup in the Northeast Atlantic Coast of the US [@MidAtlantic; @Link2010]. However, as our results show, designing these new strategies is not straightforward under the current MAFMC governance framework. Moreover, as species shift their distribution, quota allocations need to be dynamic which means they also have to account for the natural variability of the stock's distribution.

The joint management of shared stocks relies on defined boundaries (i.e., regulatory units) where each share-holder has right to fishing [@Munro]. In the northeast Pacific, for example, the joint management of Pacific halibut (*Hippoglossus stenolepis*) by Canada and the US relies on multiple regulatory areas that covers the EEZ's of both countries [@IPHC]. In the case of the MAFMC, the method for designing regulatory areas in Federal waters will have large consequences in the allocation of quota to each state (Figs. 3-5). The fishing ports approach relates to many cases around the world, where spatial fishing rights are allocated to share-holders based on historical landings [@Chile; @EU] or location of fishing grounds [@Mexico]. These methods are seen fair in that they respect the historical investment of individuals in the fishery [@REF]. However, such approach has the potential to exclude users from the fishery, like the case of Maine, New Hampshire and Connecticut (Figure 1). Moreover, because it is based on historical landings, it does not necessarily account for biomass changes across regulatory areas, nor expansion to new regulatory areas [@Pinsky2018]. The exclusion of both historic and new share-holders can result in conflict and unsustainable harvest as seen in multiple fisheries across the world and within the US [@Pinsky2018; @Pinsky2014; @ChileLoco; @Spijkers].

Designing regulatory units by expanding current state waters approach would not exclude share-holders that historically have fish less of a certain stock. In our study, by giving regulatory units to Maine, Connecticut and New Hampshire we account for the shifts in stock biomass allowing these states to benefit from the fishery as the stocks move poleward (Figures 3 and 5). However, such a definition might not be seen as fair to those share-holders that have historically invested in the fishery [@REF]. Fishers might still be willing to navigate longer distances to "follow the fish" rather than changing target species or even drooping out of the fishery [@Polcanza].

Allocating quota based on current distributions, rather than historical reference points, should improve the resilience of management plans to climate change driven shifts in stocks distribution. Modelling exercises suggest that the management plan for pacific halibut is resilient to climate change as quota is allocated to each regulatory area based on the current distribution of the stock [@Palacios-Abrantes2020; @Sumaila2020]. A combination of both historic and current distributions has also been used to manage shared stocks. The TRAC in the Gulf of Maine recommends quota for cod (*Gadus morhua*), haddock (*a*) and yellowtail flounder (*ss*) employing an algorithm that accounts 90% of stocks current distribution and 10% of historic distribution. Here, we found differences in the distribution of the stocks between the years when the quota was determined (i.e., historical period) and current values (Figures 2 and 3). This results suggest that current allocation policies might not be efficient in terms of biomass, this is specially true for black sea bass and scup.

**Uncertainty (management challenges)**

- Federal vs state waters legal perspectives
- Multiple agencies / states
- Multiple-species fisheries

**Moving forward (recommendations)**

**Conclusion**
